---
title: "Media Mix Modeling"
author: "Lunjing Yuan"
date: "January 20, 2021"
output:
  word_document: default
  html_notebook: default
  html_document: default
  pdf_document: default
---

```{r message=FALSE, warning=FALSE}

# get the required library
library(tidyverse) 
library(magrittr) 
library(caret)

cardf = read.csv("sales_ads_use.csv", header=TRUE, sep=",")


# AdStockScale() Returns the scale matrix for creating the adstock variable from the marketing per-period expenses
# Input: 
#    lambda - the discounted value
#    n - the length of the time series 

AdStockScale <- function(lambda, n) {
  
  r <- lambda^(seq_len(n)-1)
  m <- matrix(rep(r,n),nrow=n)

  z <- matrix(0,nrow=n,ncol=n)
  z[lower.tri(z,diag=TRUE)] <- m[row(m) <= (n+1-col(m))]
  z
}
```

To create the AdStock $G_t$ matrix from the current per-period marketing matrix $M_t$, all you need is to do the following. Change the code below, to try different values of $\lambda$


```{r}
MM <- as.matrix(cardf[,4:7])

AdStockScale2 <- AdStockScale(0.2,nrow(MM))
GG2 <- AdStockScale2 %*% MM

AdStockScale8 <- AdStockScale(0.8,nrow(MM))
GG8 <- AdStockScale8 %*% MM

```

Estimate the regression model with different $\lambda$ values, and choose the best model using holdout sample test. Which one do you like the best? Why?


```{r}
#using the top 30 data points as training
train_id <- 1:30

for (i in 1:9) {
# create the ad-stock variables
lambda <- i/10
GG <- AdStockScale(lambda, nrow(MM)) %*% MM 
alldata <- data.frame(cardf$sales,GG) 
train_data <- alldata[train_id,] 
test_data <- alldata[-train_id,]

# estimate the model using the training data
reg <- lm(cardf.sales~., data = train_data) 
# get the prediction using testing data 
Ypredict <- reg %>% predict(test_data)

# print results
print(paste("lambda=",lambda, " RMSE = ", RMSE(Ypredict,test_data$cardf.sales))) }

```
Based on these results, the testing data has the lowest RMSE, when λ = 0.2.

Note that in this exercise, we restricted for all the marketing variables to have the same value of λ, in real practice, this can be relaxed.
